\bf\mftitle

### 2011-02-14 doc/jInfer/basicxsd/basicxsd.tex
Documentation: BasicXSDExporter.

All settings provided by \jmodule{BasicXSDExporter} are project-wide, the preferences panel is in \code{cz.cuni.mff.ksi.jinfer.basicxsd.properties} package. As mentioned above, it is possible to set the following. 
\begin{itemize}
	\item Turn off generation of global element types. Turning off this feature is not recommended as it may cause certain problems with validity of resulting XSD. See \ref{section:problems}.
	\item Minimal number of occurrences of element to define its type globally. (Only if generation of global elements is active.)
	\item Number of spaces in output per one level of indentation.
	\item Global type name prefix. It is a string which will be inserted before a name of a type, which is derived from element's name. Can be also an empty string. (Only if generation of global elements is active.)
	\item Global type name suffix. It is a string which will be appended after a name of a type, which is derived from element's name. Can be also an empty string. (Only if generation of global elements is active.)
\end{itemize}

@@

TODO namespaces
TODO problem s vypnutym generovanim global elements
TODO problem s rovnakymi zaciatkami sekvencii

--------------------------------------------------------------------------------------

First, global elements are exported. For each element, its type is defined as a global type. TODO rio example. This is done by passing global elements to \code{processGlobalElement} method.

After global elements, others are exported by calling of method \code{processElement}, supplied by the top level element as its argument. Reference to a global element type is done simply by declaring element's name and type.

To sa deje rekurzivnym priechodom so startom v korenovom elemente. Elementy, ktorych typy su definovane globalne a elementy so vstavanymi typmi, su exportovane jednoducho, uvedenim ich mena a typu. Ostatne elementy su definovane na mieste. TODO rio priklady. Na toto je potrebne zavolat metodu \code{processElement} s korenovym elementov, ako argumentom.

Code exporting attributes is in \code{attributeToString()}. First thing this method does is to assess the domain of a particular atribute: this is a map indexed by attribute values containing number of occurences for each such attribute. Type definition of an attribute is generated in the \code{DomainUtils.getAttributeType()} method. Based on a user setting, this might decide to enumerate all possible values of this attribute using the \code{(a|b|c)} notation, otherwise it just returns \code{\#CDATA}.\\
Attribute requiredness is assessed based on \code{required} metadata presence. If an attribute is not deemed required, it might have a default value: if a certain value is prominent in the attribute domain (based on user setting again), it is declared default.

\subsection{Preferences}

All settings provided by \jmodule{BasicDTDExporter} are project-wide, the preferences panel is in \code{cz.cuni.mff.ksi.jinfer.basicdtd.properties} package. As mentioned before, it is possible to set the following. 
\begin{itemize}
	\item Maximum attribute domain size which is exported as a list of all values (\code{(a|b|c)} notation).
	\item Minimal ratio an attribute value in the domain needs to have in order to be declared default.
\end{itemize}

@@

Flow of data in this module is following.
\begin{enumerate}
	\item \code{SchemaGeneratorImpl} topologically sorts elements (rules) it got on input.
	\item For each element, relevant portion of DTD schema is generated.
	\item String representation of the schema is returned along with the information that file extension should be "dtd".
\end{enumerate}

>>>>

All settings provided by \jmodule{BasicXSDExporter} are project-wide, the preferences panel is in \code{cz.cuni.mff.ksi.jinfer.basicxsd.properties} package. As mentioned above, it is possible to set the following.

\begin{itemize}
	\item Turn off generation of global element types. Turning off this feature is not recommended as it may cause certain problems with validity of resulting XSD. See \ref{section:problems}.
	\item Minimal number of occurrences of element to define its type globally. (Only if generation of global elements is active.)
	\item Number of spaces in output per one level of indentation.
	\item Global type name prefix. It is a string which will be inserted before a name of a type, which is derived from element's name. Can be also an empty string. (Only if generation of global elements is active.)
	\item Global type name suffix. It is a string which will be appended after a name of a type, which is derived from element's name. Can be also an empty string. (Only if generation of global elements is active.)
\end{itemize}

@@

\subsection{Namespaces}

\subsubsection{Description}

This basic implementation of XSD export does not support and does not handle namespaces. That means that no namespace definitions are generated in a result, namespace definitions in input are processed as regular elements and attributes and namespace usages (namespace:element) are considered as a part of name of an element.

\subsubsection{Workaround}

To generate a valid XSD output, it is necessary to remove all namespace definitions and usages from input. If a presence of namespaces in a resulting XSD is necessary, they need to be inserted there manually.

\subsection{XSD invalidity if generation of global element types is disabled}

\subsubsection{Error conditions}

Following conditions have to be met to cause a problem.

\begin{itemize}
	\item Generation of global element types if turned off (in project properties).
	\item Input grammar contains a concatenation or an alternation regexp with several (two or more) same elements.
\end{itemize}

\subsubsection{Description}

Example of a part of generated XSD.

\begin{verbatim}
...
  <xs:sequence>
    <xs:element name="A">
      ...
    </xs:element>
    <xs:element name="A">
      ...
    </xs:element>
  <xs:sequence>
...
\end{verbatim}

In XSD, it is not allowed to define elements with the same name in one sequence (and other XSD constructs).

\subsubsection{Workaround}

Do not turn the generation of global element types off.

\subsection{"Unique Particle Attribution" problem}

\subsubsection{Error conditions}

This problem appears if input grammar contains an alternation of two or more concatenations, which begin with at least one same element. Part of a problematic XSD is for example following.

\begin{verbatim}
...
  <xs:choice>
    <xs:sequence>
      <xs:element name="A" type="TA"/>
      <xs:element name="B" type="TB"/>
    </xs:sequence>
    <xs:sequence>
      <xs:element name="A" type="TA"/>
      <xs:element name="C" type="TC"/>
    </xs:sequence>
  </choice>
...
\end{verbatim}

\subsubsection{Description}

If error conditions are met, input XMLs are not valid against a generated XSD.

@@

Flow of data in this module is following.
\begin{enumerate}
	\item \code{Preprocessor} topologically sorts elements (rules) it got on input.
	\item For each element, its occurrence count is computed.
	\item Unused elements (occurrence count equals 0) are removed.
	\item For each element, it is determined if it type will be defined as a global type or not.
	\item For each global element, its type is exported to a XSD representation by \code{GlobalElementsExporter}.
	\item Root element and recursively all remaining elements are exported to a XSD representation by \code{RootElementExporter}.
	\item String representation of the schema is returned along with the information that file extension should be "xsd".
\end{enumerate}
